Off the beaten path: \\ Gaia reveals GD-1 stars outside of the main stream

### 2018-06-20 paper/submit-aas/referee-response.txt

We don't see any correlations between astrometric issues or incompleteness and the features that we see associated with the stream.
We do, however, see density variations in the main stream track due to the scanning pattern (especially between phi1=[-80, -55]).
Given this and the above, we have added some text to Section 3.1 to describe this, but have decided not to add the attached figure to the text because we have instead included a new figure to address a later point (see below).

Lastly, we have decided not to include further zoom-in panels, as our figures already contain many panels, but we have tightened the axis ranges on the proper motion panels of (what is now) Figure 5 and hope that this makes the distribution in proper-motion space clearer.

> The authors need to include a measurement of the proper motion in the
> equatorial reference frame, for example, at the location of the highest
> density of the stream. This is clearly missing from the paper, and would make
> the paper useful to other readers.

We have added this quantity to the paper in Section 2:
---
At this location, between $\phi_1 \in (-18, -10)^\circ$, we find that the mean equatorial sky position of the stream stars is $(\alpha, \delta) = (177.01, 53.99)^\circ$ and the mean proper motion is $(\mu_\alpha \, \cos\delta, \mu_\delta) = (-7.78, -7.85) \pm 0.03~\masyr$, taking into account the covariance matrix for proper motions provided with the \gaia\ data.
---

### 2018-06-20 paper/submit-aas/referee-response.txt

We thank the referee for these suggestions and agree that it is useful and important to discuss the quality of the astrometric solution in this region of the sky.
We have attached a figure to this report that shows maps of (1) the density of 2-parameters solutions in this region, (2) the unit weight error (as defined in Lindegren+2018, C.1), and (3) the SFD dust extinction around GD-1.
We have also computed the source density at several locations along the stream to assess whether source crowding could impact the completeness (as in Arenou+2018).
We find that, at each location along the stream between phi1 = (-50, -10) the source density is never larger than ~8000 sources/deg^2 (estimated using the 2-parameter and 5-parameter sources), and is far below the densities where crowding becomes a problem.

We don't see any correlations between astrometric issues or incompleteness and the features that we see associated with the stream.
We do, however, see density variations in the main stream track due to the scanning pattern (especially between phi1=[-80, -55]).
Given this and the above, we have added some text to Section 3.1 to note this, but have decided not to add the attached figure to the text because we have instead included a new figure to address a later point (see below).

Lastly, we have decided not to include further zoom-in panels, as our figures already contain many panels, but we have tightened the axis ranges on the proper motion panels of (what is now) Figure 5.
We hope that this makes the distribution in proper-motion space clearer.

### 2018-06-20 paper/gd1.tex
software citations

We retrieve data along the previously-identified track of the GD-1 stream
from the \gaia\ science archive\footnote{\url{https://gea.esac.esa.int/}} by
selecting sources with small parallax, $\varpi < 1~\textrm{mas}$ in fields along the GD-1 stream.
% , and select
% batches of stars that have small latitude in the heliocentric GD-1 coordinate
% system, $(\phi_1, \phi_2)$, defined in \cite{Koposov:2010}.
We convert the equatorial sky coordinates and proper motions %of the parallax-selected stars
to the GD-1 coordinate system \citep[$\phi_1, \phi_2$,][]{Koposov:2010}, and correct the proper motions for solar reflex
motion by assuming that stars at a given stream longitude, $\phi_1$, have a
distance given by $d(\phi_1) = (0.05 \, \phi_1 + 10)~\textrm{kpc}$;
we assume a solar velocity $\bs{v}_\odot = (11.1, 232.24, 7.25)~\kms$
(\citealt{Schonrich:2010, Bovy:2015}).
\changes{In detail, we project the solar velocity vector onto the tangent space at each star's sky position, convert to angular motion using the assumed distance, and subtract the solar component from each star's proper motion (this is done using the \package{astropy} and \package{gala} packages).}
\figurename~\ref{fig:selection} (top right) shows stars with $|\phi_2| <
1^\circ$ in solar-reflex-corrected proper motion components in the GD-1 system,
$\mu_{\phi_1, \odot}$ and $\mu_{\phi_2, \odot}$.
GD-1 members are visible as the overdensity around $(\mu_{\phi_1, \odot},
\mu_{\phi_2, \odot}) \approx (-8, 0)~\masyr$.

>>>>

We retrieve data along the previously-identified track of the GD-1 stream
from the \gaia\ science archive\footnote{\url{https://gea.esac.esa.int/}} by
selecting sources with small parallax, $\varpi < 1~\textrm{mas}$ in fields along the GD-1 stream.
% , and select
% batches of stars that have small latitude in the heliocentric GD-1 coordinate
% system, $(\phi_1, \phi_2)$, defined in \cite{Koposov:2010}.
We convert the equatorial sky coordinates and proper motions %of the parallax-selected stars
to the GD-1 coordinate system \citep[$\phi_1, \phi_2$,][]{Koposov:2010}, and correct the proper motions for solar reflex
motion by assuming that stars at a given stream longitude, $\phi_1$, have a
distance given by $d(\phi_1) = (0.05 \, \phi_1 + 10)~\textrm{kpc}$;
we assume a solar velocity $\bs{v}_\odot = (11.1, 232.24, 7.25)~\kms$
(\citealt{Schonrich:2010, Bovy:2015}).
\changes{In detail, we project the solar velocity vector onto the tangent space at each star's sky position, convert to angular motion using the assumed distance, and subtract the solar component from each star's proper motion (this is done using the \package{astropy} \citep{astropy} and \package{gala} \citep{gala} packages).}
\figurename~\ref{fig:selection} (top right) shows stars with $|\phi_2| <
1^\circ$ in solar-reflex-corrected proper motion components in the GD-1 system,
$\mu_{\phi_1, \odot}$ and $\mu_{\phi_2, \odot}$.
GD-1 members are visible as the overdensity around $(\mu_{\phi_1, \odot},
\mu_{\phi_2, \odot}) \approx (-8, 0)~\masyr$.
